Streaking and Clutching in Round-Based First Person Shooter Esports

### 2022-07-17 presentation.qmd
more val processing, prefix cod vars

---
title: "Streaking and Clutching in Round-Based First Person Shooter Esports"
author: "Tony ElHabr"
format:
  revealjs:
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

rounds <- qs::qread('cod_rounds.qs')
```

@@

::: fragment
-   Plant/Defuse in Valorant
    -   Valorant professional league, updated through January 1, 2022
    -   5v5
    -   First to 13 round wins
    -   One team plays offense in 12 consecutive rounds before swapping sides
    -   Must win by 2 rounds
:::

@@

```{r}
#| label: descriptive-cod-snd-stats
#| eval: true
n_series <- rounds |> 
  distinct(year, sheet, series) |> 
  nrow() |> 
  scales::comma()

n_rounds <- rounds |> 
  distinct(year, sheet, series, round) |> 
  nrow() |> 
  scales::comma()

o_win_prop_const <- rounds |> 
  filter(is_offense) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop)
o_win_pct <- scales::percent(o_win_prop_const, accuracy = 1)

plant_pct <- rounds |>
  filter(is_offense) |> 
  count(makes_plant = !is.na(plant)) |> 
  mutate(prop = n / sum(n)) |> 
  filter(makes_plant) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

o_plant_and_win_pct <- rounds |> 
  filter(is_offense, !is.na(plant)) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

o_first_blood_win_pct <- rounds |> 
  filter(!is.na(earned_fb)) |> 
  filter(earned_fb) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)
```

::: fragment
-   Total series: **`r n_series`**
:::

::: fragment
-   Total rounds: **`r n_rounds`**
:::

::: fragment
-   Offensive plant %: **`r plant_pct`**
:::

::: fragment
-   Offensive win %: **`r o_win_pct`**
:::

::: fragment
-   Win % when planting: **`r o_plant_and_win_pct`**
:::

::: fragment
-   Win % when earning first blood: **`r o_first_blood_win_pct`**
:::

@@

```{r}
#| label: cod_heatmap_prep
n_rounds_side <- rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

o_win_prop <- n_rounds_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!o_win_prop_const
  )

e_o_win_series_prop <- crossing(
  pre_cumu_w = 0:5,
  pre_cumu_l = 0:5
) |> 
  mutate(
    n_remain_max = 11 - (pre_cumu_w + pre_cumu_l),
    n_remain_w = 6 - pre_cumu_w,
    ## can i account for offense in this by rotating?
    e_win_series_prop = map2_dbl(n_remain_w, n_remain_max, ~sum(dbinom(..1:..2, ..2, 0.5)))
  )

o_win_prop <- o_win_prop |> 
  inner_join(
    e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  )

@@

heatmap_seq_labeller <- function(.data, .label) {
  sprintf(
    '%s\n(%s/%s)', 
    scales::percent(.data[[.label]], accuracy = 1), 
    scales::comma(.data$win_round), 
    scales::comma(.data$n)
  )
}

heatmap_text_layer <- function(..., .op, .data, .label, .color, .threshold, .labeller) {
  list(
    ...,
    geom_text(
      data = .data |> filter(`.op`(.data[[.label]], .threshold)),
      # family = font,
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        x = pre_cumu_w + 0.5, 
        y = pre_cumu_l + 0.5,
        label = .labeller(.data = .data, .label = .label)
      )
    )
  )
}

heatmap_text_layers <- function(..., .f) {
  list(
    ...,
    .f(.emphasize = 'low'),
    .f(.emphasize = 'high')
  )
}
```

```{r}
#| label: p_o_round_win_prop
o_round_win_text_layer <- function(..., .emphasize = 'low') {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_round_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_round_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(.f = o_round_win_text_layer) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_o_round_win_prop
save_for_slide(p_o_round_win_prop)
```

@@

![](o_round_win_prop.png){fig-align="center"}

## How do the empirical round results compare to expectation?

```{r}
#| label: p_o_round_win_prop_over_expected
heatmap_div_labeller <- function(.data, .label) {
  sprintf(
    '%+.0f\n(%s/%s)', 
    100 * .data[[.label]], 
    scales::comma(.data$win_round), 
    scales::comma(.data$n)
  )
}

o_round_win_over_expected_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- 'white'
  } else {
    op <- `>=`
    .color <- blackish_background
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'diff_win_round_prop',
      .color = .color,
      .threshold = -0.1,
      .labeller = heatmap_div_labeller
    )
  )
}

@@

max_diff_win_round_prop <- max(abs(o_win_prop$diff_win_round_prop))
p_o_round_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(.f = o_round_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_round_prop, 
      +max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_o_round_win_prop_over_expected
save_for_slide(p_o_round_win_prop_over_expected)
```

@@

![](o_round_win_prop_over_expected.png){fig-align="center"}

@@

```{r}
#| label: p_o_series_win_prop
add_o_series_win_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_series_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_for_slide(p_o_series_win_prop)
```

@@

![](o_series_win_prop.png){fig-align="center"}

## How do the empirical series results compare to expectation?

```{r}
#| label: p_o_series_win_prop_over_expected
add_o_series_win_over_expected_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'diff_win_series_prop',
      .color = .color,
      .threshold = 0.08,
      .labeller = heatmap_div_labeller
    )
  )
}

max_diff_win_series_prop <- max(abs(o_win_prop$diff_win_series_prop))
p_o_series_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_series_prop, 
      +max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_o_series_win_prop_over_expected
save_for_slide(p_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: o_win_r1_vs_post_round_pro
calculate_o_round_win_prop <- function(df) {
  df |> 
    count(year, game, map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      year,
      game,
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

@@

summarize_o_round_win_prop <- function(.round) {

  specified_round <- rounds |> 
    filter(round == .round) |> 
    calculate_o_round_win_prop()

  post_rounds <- rounds |> 
    filter(round > .round) |> 
    calculate_o_round_win_prop()

@@

plot_round_vs_post_round_slope_chart <- function(.round) {
  o_win_rx_prop <- summarize_o_round_win_prop(.round)
  x1_label <- sprintf('Round %d<br/>o Win %%', .round)
  x2_label <- sprintf('Post-Round %d<br/>o Win %%', .round)

@@

res_r1 <- plot_round_vs_post_round_slope_chart(1)

p_o_win_r1_vs_post_round_prop <- res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  )
p_o_win_r1_vs_post_round_prop
save_for_slide(p_o_win_r1_vs_post_round_prop)
```

![](o_win_r1_vs_post_round_prop.png){fig-align="center"}

@@

## Are offensive win %'s different in round 2?

```{r}
#| label: o_win_r2_vs_post_round_pro
res_r2 <- plot_round_vs_post_round_slope_chart(2)
res_r2$plot
save_for_slide(
  res_r2$plot,
  filename = 'o_win_r2_vs_post_round_prop.png'
)
```

![](o_win_r2_vs_post_round_prop.png){fig-align="center"}

::: notes
-   no trend
:::

@@

```{r}
#| label: p_series_outcome_prop_over_expected
# https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422
prob_of_series_lasting_n_games <- function(n, p = 0.5, max_rounds = 11) {
  s <- ceiling(max_rounds / 2)
  (factorial(n - 1) / (factorial(s - 1) * factorial(n - s))) * (p^s * (1 - p)^(n - s) + p^(n - s) * (1 - p)^s)
}

@@

actual_series_streaks <- rounds |> 
  filter(win_series) |> 
  mutate(across(win_round, as.integer)) |> 
  group_by(year, event, series) |> 
  summarize(
    wins = max(cumu_w),
    losses = max(cumu_l),
    ws = paste0(win_round, collapse = '')
  ) |> 
  ungroup() |> 
  mutate(n_rounds = wins + losses) |> 
  unite(
    record, wins, losses, sep = '-'
  ) |> 
  count(n_rounds, record, ws, sort = TRUE) |> 
  mutate(prop = n / sum(n))

possibilities <- tibble(
  n_rounds = 6L:11L
) |> 
  mutate(
    series_prop = map_dbl(n_rounds, prob_of_series_lasting_n_games),
    possibilities = map2_dbl(6, n_rounds - 6, possibilities_of_outcomes)
  ) |> 
  transmute(
    n_rounds,
    series_prop,
    record = sprintf('6-%d', n_rounds - 6),
    prop = series_prop / possibilities
  )

series_streaks <- inner_join(
  actual_series_streaks |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  possibilities |> 
    rename_with(~sprintf('%s_expected', .x), prop),
  by = 'n_rounds'
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected
  ) |> 
  arrange(desc(abs(prop_diff)))
# series_streaks |> filter(is.na(n_actual))

## 6-0, 6-1, and some combos of 6-5 occur more often than expected
series_streak_props <- series_streaks |> 
  filter(!is.na(n_actual)) |> 
  mutate(
    prop_diff = prop_actual - prop_expected,
    total_actual = sum(n_actual),
    n_expected = round(prop_expected * total_actual),
    p = vectorized_prop_test(n_actual, total_actual, n_expected, total_actual)
  ) |> 
  select(-total_actual) |> 
  unnest_wider(p) |> 
  arrange(p.value)

>>>>

---
title: "Streaking and Clutching in Round-Based First Person Shooter Esports"
author: "Tony ElHabr"
format:
  revealjs:
    slide-number: true
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

cod_rounds <- qs::qread('cod_rounds.qs')
```

@@

::: fragment
-   Plant/Defuse in Valorant
-   Valorant professional league, updated through January 1, 2022
-   5v5
-   First to 13 round wins
-   One team plays offense in 12 consecutive cod_rounds before swapping sides
-   Must win by 2 cod_rounds
:::

@@

```{r}
#| label: descriptive-cod-snd-stats
#| eval: true
cod_n_series <- cod_rounds |> 
  distinct(year, sheet, series) |> 
  nrow() |> 
  scales::comma()

cod_n_rounds <- cod_rounds |> 
  distinct(year, sheet, series, round) |> 
  nrow() |> 
  scales::comma()

cod_o_win_prop_const <- cod_rounds |> 
  filter(is_offense) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop)
cod_o_win_pct <- scales::percent(cod_o_win_prop_const, accuracy = 1)

cod_plant_pct <- cod_rounds |>
  filter(is_offense) |> 
  count(makes_plant = !is.na(plant)) |> 
  mutate(prop = n / sum(n)) |> 
  filter(makes_plant) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

cod_o_plant_and_win_pct <- cod_rounds |> 
  filter(is_offense, !is.na(plant)) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

cod_o_first_blood_win_pct <- cod_rounds |> 
  filter(!is.na(earned_fb)) |> 
  filter(earned_fb) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)
```

::: fragment
-   Total series: **`r cod_n_series`**
:::

::: fragment
-   Total cod_rounds: **`r cod_n_rounds`**
:::

::: fragment
-   Offensive plant %: **`r cod_plant_pct`**
:::

::: fragment
-   Offensive win %: **`r cod_o_win_pct`**
:::

::: fragment
-   Win % when planting: **`r cod_o_plant_and_win_pct`**
:::

::: fragment
-   Win % when earning first blood: **`r cod_o_first_blood_win_pct`**
:::

@@

```{r}
#| label: cod_heatmap_prep
cod_round_and_series_win_prop_by_side <- cod_rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

cod_o_win_prop <- cod_round_and_series_win_prop_by_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!cod_o_win_prop_const
  )

cod_e_o_win_series_prop <- crossing(
  pre_cumu_w = 0:5,
  pre_cumu_l = 0:5
) |> 
  mutate(
    n_remain_max = 11 - (pre_cumu_w + pre_cumu_l),
    n_remain_w = 6 - pre_cumu_w,
    ## can i account for offense in this by rotating?
    e_win_series_prop = map2_dbl(n_remain_w, n_remain_max, ~sum(dbinom(..1:..2, ..2, 0.5)))
  )

cod_o_win_prop <- cod_o_win_prop |> 
  inner_join(
    cod_e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  )

# round_and_series_win_prop <- cod_rounds |> 
#   group_by(pre_cumu_w, pre_cumu_l) |> 
#   summarize(
#     n = n(),
#     across(c(win_round, win_series), sum)
#   ) |> 
#   ungroup() |> 
#   mutate(
#     win_round_prop = win_round / n,
#     win_series_prop = win_series / n
#   )
# 
# pluck_round_win_prop <- function(w = 0, l = 0) {
#   round_and_series_win_prop |> 
#     filter(pre_cumu_w == w, pre_cumu_l == l) |> 
#     pull(win_round_prop)
# }
# 
# compute_conditional_probs <- function(
    #     p00 = 0.5,
#     p10 = p00,
#     p20 = p00,
#     p30 = p00,
#     p40 = p00,
#     p50 = p00,
#     p11 = p00, p21 = p00, p31 = p00, p41 = p00, p51 = p00,
#     p52 = p00, p53 = p00, p54 = p00
# ) {
#   pr6 <- p10 * p20 * p30 * p40 * p50
#   # pr7 <- p51 * (p10 * p20 * p30 * p40 * (1 - p50) + p10 * p20 * p30 * (1 - p40) * p41 + p10 * p20 * (1 - p30) * p41 + p10 * (1 - p20) * p31 * p41 + (1 - p10) * p21 * p31 * p41)
#   pr7 <- (
#     (p00 * p10 * p20 * p30 * p40 * (1 - p50) * p51) +   ## aaaaaba
#     (p00 * p10 * p20 * p30 * (1 - p40) * p41 * p51) + ## aaaabaa
#     (p00 * p10 * p20 * (1 - p30) * p31 * p41 * p51) + ## aaabaaa
#     (p00 * p10 * (1 - p20) * p21 * p31 * p41 * p51) + ## aabaaaa
#     (p00 * (1 - p10) * p11 * p21 * p31 * p41 * p51) + ## abaaaaa
#     ((1 - p00) * (1 - p10) * p11 * p21 * p31 * p41 * p51) + ## baaaaaa
#     
#     ((1 - p00) * p10 * p20 * p30 * p40 * (1 - p50) * p51) + ## bbbbbab
#     ((1 - p00) * p10 * p20 * p30 * (1 - p40) * p41 * p51) + ## bbbbabb
#     ((1 - p00) * p10 * p20 * (1 - p30) * p31 * p41 * p51) + ## bbbabbb
#     ((1 - p00) * p10 * (1 - p20) * p21 * p31 * p41 * p51) + ## bbabbbb
#     ((1 - p00) * (1 - p10) * (1 - p11) * p21 * p31 * p41 * p51) + ## babbbbb
#     (p00 * (1 - p10) * (1 - p11) * p21 * p31 * p41 * p51) ## abbbbbb
# )
#   
#   ## NOTE: pr8 and pr9 are broken. although we can use the simple subtraction of other probabilities for pr10 and pr11 like pr6 and pr7 in a 7 game series, they aren't valid for pr8 and pr9 in an 11 game series.
#   pr8 <- (1 - pr6 - pr7) * p52
#   pr9 <- (1 - pr6 - pr7 - pr8) * p53
#   pr10 <- (1 - pr6 - pr7 - pr8 - pr9) * p54
#   pr11 <- 1 - pr6 - pr7 - pr8 - pr9 - pr10
#   tibble(
#     n_rounds = 6L:11L,
#     p = c(pr6, pr7, pr8, pr9, pr10, pr11)
#   )
# }
# 
# ## pr6 and pr7 should match up with what we see later with `possibilities`
# conditional_series_probs <- compute_conditional_probs(
#   p00 = pluck_round_win_prop(0, 0),
#   p10 = pluck_round_win_prop(1, 0),
#   p20 = pluck_round_win_prop(2, 0),
#   p30 = pluck_round_win_prop(3, 0),
#   p40 = pluck_round_win_prop(4, 0),
#   p50 = pluck_round_win_prop(5, 0),
#   p11 = pluck_round_win_prop(1, 1),
#   p21 = pluck_round_win_prop(2, 1),
#   p31 = pluck_round_win_prop(3, 1),
#   p41 = pluck_round_win_prop(4, 1),
#   p51 = pluck_round_win_prop(5, 1),
#   p52 = pluck_round_win_prop(5, 2),
#   p53 = pluck_round_win_prop(5, 3),
#   p54 = pluck_round_win_prop(5, 4)
# )

@@

heatmap_seq_labeller <- function(.df, .label, .num) {
  sprintf(
    '%s\n(%s/%s)',
    scales::percent(.df[[.label]], accuracy = 1),
    scales::comma(.df[[.num]]),
    scales::comma(.df$n)
  )
}

heatmap_text_layer <- function(..., .op, .df, .label, .num, .color, .threshold, .labeller) {
  .df_filt <- .df |> filter(`.op`(.df[[.label]], .threshold))
  if(nrow(.df_filt) == 0) {
    return(list(...))
  }
  list(
    ...,
    geom_text(
      data = .df_filt,
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        label = .labeller(.df = .df_filt, .label = .label, .num = .num),
        x = .data$pre_cumu_w + 0.5, 
        y = .data$pre_cumu_l + 0.5
      )
    )
  )
}

heatmap_text_layers <- function(
    ..., 
    .df,
    .threshold,
    .labeller,
    .f
) {
  list(
    ...,
    .f(.df = .df, .threshold = .threshold, .labeller = .labeller, .emphasize = 'low'),
    .f(.df = .df, .threshold = .threshold, .labeller = .labeller, .emphasize = 'high')
  )
}

o_round_win_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'win_round_prop',
      .num = 'win_round',
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

p_cod_o_round_win_prop <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop,
    .f = o_round_win_text_layer,
    .threshold = 0.52,
    .labeller = heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_cod_o_round_win_prop
save_for_slide(p_cod_o_round_win_prop)
```

@@

![](cod_o_round_win_prop.png){fig-align="center"}

```{r}
#| label: p_cod_o_round_win_prop_over_expected
# I have no idea why I need to add .num here... keeps erroring otherwise
heatmap_div_labeller <- function(.df, .label, .num) {
  sprintf(
    '%+.1f\n(%s)', 
    100 * .df[[.label]], 
    scales::comma(.df$n)
  )
}

o_round_win_over_expected_text_layer <- function(
    ..., 
    .df, 
    .threshold,
    .labeller,
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- 'white'
  } else {
    op <- `>=`
    .color <- blackish_background
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'diff_win_round_prop',
      .num = 1,
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

@@

cod_max_diff_win_round_prop <- max(abs(cod_o_win_prop$diff_win_round_prop))
p_cod_o_round_win_prop_over_expected <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_round_win_over_expected_text_layer,
    .threshold = -0.1,
    .labeller = heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -cod_max_diff_win_round_prop, 
      +cod_max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_cod_o_round_win_prop_over_expected
save_for_slide(p_cod_o_round_win_prop_over_expected)
```

@@

![](cod_o_round_win_prop_over_expected.png){fig-align="center"}

@@

```{r}
#| label: p_cod_o_series_win_prop
o_series_win_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'win_series_prop',
      .num = 'win_series',
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

p_cod_o_series_win_prop <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_series_win_text_layer,
    .threshold = 0.8,
    .labeller = heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_cod_o_series_win_prop
save_for_slide(p_cod_o_series_win_prop)
```

@@

![](cod_o_series_win_prop.png){fig-align="center"}

```{r}
#| label: p_cod_o_series_win_prop_over_expected
o_series_win_over_expected_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'diff_win_series_prop',
      .num = NULL,
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

cod_max_diff_win_series_prop <- max(abs(cod_o_win_prop$diff_win_series_prop))
p_cod_o_series_win_prop_over_expected <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.08,
    .labeller = heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -cod_max_diff_win_series_prop, 
      +cod_max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_cod_o_series_win_prop_over_expected
save_for_slide(p_cod_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: p_cod_o_win_r1_vs_post_round_prop
calculate_cod_o_round_win_prop <- function(df) {
  df |> 
    count(year, game, map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      year,
      game,
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

@@

summarize_cod_o_round_win_prop <- function(.round) {

  specified_round <- cod_rounds |> 
    filter(round == .round) |> 
    calculate_cod_o_round_win_prop()

  post_rounds <- cod_rounds |> 
    filter(round > .round) |> 
    calculate_cod_o_round_win_prop()

@@

plot_cod_round_vs_post_round_slope_chart <- function(.round) {
  o_win_rx_prop <- summarize_cod_o_round_win_prop(.round)
  x1_label <- sprintf('Round %d<br/>offensive win %%', .round)
  x2_label <- sprintf('Post-round %d<br/>offensive win %%', .round)

@@

cod_res_r1 <- plot_cod_round_vs_post_round_slope_chart(1)

p_cod_o_win_r1_vs_post_round_prop <- cod_res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * cod_res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * cod_res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  )
p_cod_o_win_r1_vs_post_round_prop
save_for_slide(p_cod_o_win_r1_vs_post_round_prop)
```

![](cod_o_win_r1_vs_post_round_prop.png){fig-align="center"}

@@

@@

```{r}
#| label: p_cod_series_outcome_prop_over_expected
# https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422
prob_of_series_lasting_n_games <- function(n, p = 0.5, max_rounds = 11) {
  s <- ceiling(max_rounds / 2)
  (factorial(n - 1) / (factorial(s - 1) * factorial(n - s))) * (p^s * (1 - p)^(n - s) + p^(n - s) * (1 - p)^s)
}

@@

cod_actual_series_streaks <- cod_rounds |> 
  filter(win_series) |> 
  mutate(across(win_round, as.integer)) |> 
  group_by(year, event, series) |> 
  summarize(
    wins = max(cumu_w),
    losses = max(cumu_l),
    ws = paste0(win_round, collapse = '')
  ) |> 
  ungroup() |> 
  mutate(n_rounds = wins + losses) |> 
  unite(
    record, wins, losses, sep = '-'
  ) |> 
  count(n_rounds, record, ws, sort = TRUE) |> 
  mutate(prop = n / sum(n))

cod_possibilities <- tibble(
  n_rounds = 6L:11L
) |> 
  mutate(
    series_prop = map_dbl(n_rounds, prob_of_series_lasting_n_games),
    possibilities = map2_dbl(6, n_rounds - 6, possibilities_of_outcomes)
  ) |> 
  transmute(
    n_rounds,
    series_prop,
    record = sprintf('6-%d', n_rounds - 6),
    prop = series_prop / possibilities
  )

cod_series_streaks <- inner_join(
  cod_actual_series_streaks |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  cod_possibilities |> 
    rename_with(~sprintf('%s_expected', .x), prop),
  by = 'n_rounds'
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected
  ) |> 
  arrange(desc(abs(prop_diff)))
# series_streaks |> filter(is.na(n_actual))

## 6-0, 6-1, and some combos of 6-5 occur more often than expected
cod_series_streak_props <- cod_series_streaks |> 
  filter(!is.na(n_actual)) |> 
  mutate(
    prop_diff = prop_actual - prop_expected,
    total_actual = sum(n_actual),
    n_expected = round(prop_expected * total_actual),
    p = vectorized_prop_test(n_actual, total_actual, n_expected, total_actual)
  ) |> 
  select(-total_actual) |> 
  unnest_wider(p) |> 
  arrange(p.value)
